---
title: "Programa de Prevención y Control de Arbovorisis | 2026 | Guanajuato"
format: 
  dashboard:
    nav-buttons: [twitter, github, facebook]
    logo: https://framework-gb.cdn.gob.mx/gobmx/img/logo_blanco.svg
    favicon: https://framework-gb.cdn.gob.mx/gobmx/img/logo_blanco.svg
theme: custom.scss
---

# [**Hotspots de Dengue**]{style="font-size:2pt"}
 
::: panel-tabset

### [**León de Aldama**]{style="color:#990000"}

<html>
<head>
```{=html}
<style>
* {
  box-sizing: border-box;
}

body {
  margin: 0;
  padding: 0;
  width: 100%;
  height: auto;
}

.wrapper {
  display: grid;
  gap: 1px;
  background-color: #772953;
  width: 100%;
  padding: 1px;

  /* Desktop: layout original */
  grid-template-columns: 1fr 1fr 1fr 1.8fr;
  grid-template-rows: auto auto auto auto;
}

.box {
  background-color: #772953;
  color: #772953;
  border-radius: 4px;
  padding: 4px;
  font-size: 100%;
  min-height: 80px;
}

/* Desktop grande (>= 1024px): layout original */
@media (min-width: 1024px) {
  .wrapper {
    display: grid;
    grid-gap: 1px;
    grid-template-columns: 1fr 1fr 1fr 1.8fr;
    grid-template-rows: calc(100vh/4.5) calc(100vh/4.5) calc(100vh/4.5) calc(100vh/5);
    background-color: #772953;
    color: "#772953";
    width: 100%;
    height: 100vh;
    box-sizing: border-box;
}
  .a { grid-column: 1 / 4; grid-row: 1 / 4; }
  .b { grid-column: 4;     grid-row: 1 / 3; }
  .c { grid-column: 4;     grid-row: 3 / 5; }
  .d { grid-column: 3;     grid-row: 4;     }
  .e { grid-column: 2;     grid-row: 4;     }
  .f { grid-column: 1;     grid-row: 4;     }
}

/* Tablet iPad (768px – 1023px): mapa arriba, gráficas abajo */
@media (min-width: 768px) and (max-width: 1023px) {
  .wrapper {
    grid-template-columns: 1fr 1fr;
    grid-template-rows: auto;
  }
  .a { grid-column: 1 / 3; grid-row: 1; min-height: 400px; }
  .b { grid-column: 1;     grid-row: 2; min-height: 280px; }
  .c { grid-column: 2;     grid-row: 2; min-height: 280px; }
  .d { grid-column: 1;     grid-row: 3; min-height: 120px; }
  .e { grid-column: 2;     grid-row: 3; min-height: 120px; }
  .f { grid-column: 1 / 3; grid-row: 4; min-height: 120px; }
}

/* Smartphone (< 768px): columna única */
@media (max-width: 767px) {
  .wrapper {
    grid-template-columns: 1fr;
    grid-template-rows: auto;
  }
  .a { grid-column: 1; grid-row: 1; min-height: 300px; }
  .b { grid-column: 1; grid-row: 2; min-height: 250px; }
  .c { grid-column: 1; grid-row: 3; min-height: 250px; }
  .d { grid-column: 1; grid-row: 4; min-height: 110px; }
  .e { grid-column: 1; grid-row: 5; min-height: 110px; }
  .f { grid-column: 1; grid-row: 6; min-height: 110px; }
}
</style>
```
</head>
<body>
::: wrapper
::: {.box .a}
```{r hotspots_dengue_leon}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "11", 
                                 locality = "León de Aldama")

# Step 2. Extract the hotspots ####
hotspots <- hotspotsdata::hotspots_hex[loc, ]
# Step 3. map
pal <- rcartocolor::carto_pal(n = max(hotspots$intensity_gi), name = "OrYel")
pal_fn <- grDevices::colorRampPalette(pal)
mapgl::maplibre_view(hotspots,
                     column = "intensity_gi",
                     palette = pal_fn) |>
    mapgl::add_continuous_legend(legend_title = "Intensidad",
                          values = seq(from = 1, 
                                       to = max(hotspots$intensity_gi),
                                       by = 1),
                          colors = pal,
                          position = "top-left",
                          width = "250px",
                          draggable = TRUE,
                          style = mapgl::legend_style(element_border_width =0,
                                                      shadow_size = 0,
                                                      shadow = FALSE),
                          interactive = TRUE) |>
    mapgl::add_fullscreen_control(position = "top-left") |> 
    mapgl::add_navigation_control() |>
    mapgl::add_globe_control() |>
    mapgl::add_scale_control()

```
:::

::: {.box .b}
```{r forest_plot_leon, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_leon,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto) +
                  ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                                 axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")))  |> 
  plotly::layout(showlegend = FALSE)






```
:::

::: {.box .d}
```{r gauge_blocks_leon,fig.align = "center", out.height='90%', out.width = '90%'}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_leon, fig.align = "center", out.height='90%', out.width = '90%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_leon, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::
</body>
</html>


### [**Guanajuato**]{style="color:#990000"}

<html>
<head>

</head>
<body>
::: wrapper
::: {.box .a}
```{r hotspots_dengue_guanajuato}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "11", 
                                 locality = "Guanajuato")

# Step 2. Extract the hotspots ####
#hotspots <- hotspotsdata::den_hotspots[loc, ]

hotspots <- hotspotsdata::hotspots_hex[loc, ]

# Step 3. map
pal <- rcartocolor::carto_pal(n = max(hotspots$intensity_gi), name = "OrYel")
pal_fn <- grDevices::colorRampPalette(pal)
mapgl::maplibre_view(hotspots,
                     column = "intensity_gi",
                     palette = pal_fn) |>
    mapgl::add_continuous_legend(legend_title = "Intensidad",
                          values = seq(from = 1, 
                                       to = max(hotspots$intensity_gi),
                                       by = 1),
                          colors = pal,
                          position = "top-left",
                          width = "250px",
                          draggable = TRUE,
                          style = mapgl::legend_style(element_border_width =0,
                                                      shadow_size = 0,
                                                      shadow = FALSE),
                          interactive = TRUE) |>
    mapgl::add_fullscreen_control(position = "top-left") |> 
    mapgl::add_navigation_control() |>
    mapgl::add_globe_control() |>
    mapgl::add_scale_control()

```
:::

::: {.box .b}
```{r forest_plot_guanajuato, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_guanajuato,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto) +
                  ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                                 axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")))  |> 
  plotly::layout(showlegend = FALSE)






```
:::

::: {.box .d}
```{r gauge_blocks_guanajuato,fig.align = "center", out.height='90%', out.width = '90%'}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_guanajuato, fig.align = "center", out.height='90%', out.width = '90%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_guanajuato, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::
</body>
</html>



### [**Celaya**]{style="color:#990000"}

<html>
<head>

</head>
<body>
::: wrapper
::: {.box .a}
```{r hotspots_dengue_celaya}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "11", 
                                 locality = "Celaya")

# Step 2. Extract the hotspots ####
#hotspots <- hotspotsdata::den_hotspots[loc, ]

hotspots <- hotspotsdata::hotspots_hex[loc, ]

# Step 3. map
pal <- rcartocolor::carto_pal(n = max(hotspots$intensity_gi), name = "OrYel")
pal_fn <- grDevices::colorRampPalette(pal)
mapgl::maplibre_view(hotspots,
                     column = "intensity_gi",
                     palette = pal_fn) |>
    mapgl::add_continuous_legend(legend_title = "Intensidad",
                          values = seq(from = 1, 
                                       to = max(hotspots$intensity_gi),
                                       by = 1),
                          colors = pal,
                          position = "top-left",
                          width = "250px",
                          draggable = TRUE,
                          style = mapgl::legend_style(element_border_width =0,
                                                      shadow_size = 0,
                                                      shadow = FALSE),
                          interactive = TRUE) |>
    mapgl::add_fullscreen_control(position = "top-left") |> 
    mapgl::add_navigation_control() |>
    mapgl::add_globe_control() |>
    mapgl::add_scale_control()

```
:::

::: {.box .b}
```{r forest_plot_celaya, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_celaya,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto) +
                  ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                                 axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")))  |> 
  plotly::layout(showlegend = FALSE)






```
:::

::: {.box .d}
```{r gauge_blocks_celaya,fig.align = "center", out.height='90%', out.width = '90%'}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_celaya, fig.align = "center", out.height='90%', out.width = '90%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_celaya, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::
</body>
</html>



### [**Irapuato**]{style="color:#990000"}

<html>
<head>

</head>
<body>
::: wrapper
::: {.box .a}
```{r hotspots_dengue_irapuato}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "11", 
                                 locality = "Irapuato")

# Step 2. Extract the hotspots ####
#hotspots <- hotspotsdata::den_hotspots[loc, ]

hotspots <- hotspotsdata::hotspots_hex[loc, ]

# Step 3. map
pal <- rcartocolor::carto_pal(n = max(hotspots$intensity_gi), name = "OrYel")
pal_fn <- grDevices::colorRampPalette(pal)
mapgl::maplibre_view(hotspots,
                     column = "intensity_gi",
                     palette = pal_fn) |>
    mapgl::add_continuous_legend(legend_title = "Intensidad",
                          values = seq(from = 1, 
                                       to = max(hotspots$intensity_gi),
                                       by = 1),
                          colors = pal,
                          position = "top-left",
                          width = "250px",
                          draggable = TRUE,
                          style = mapgl::legend_style(element_border_width =0,
                                                      shadow_size = 0,
                                                      shadow = FALSE),
                          interactive = TRUE) |>
    mapgl::add_fullscreen_control(position = "top-left") |> 
    mapgl::add_navigation_control() |>
    mapgl::add_globe_control() |>
    mapgl::add_scale_control()

```
:::

::: {.box .b}
```{r forest_plot_irapuato, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_irapuato,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto) +
                  ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                                 axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")))  |> 
  plotly::layout(showlegend = FALSE)






```
:::

::: {.box .d}
```{r gauge_blocks_irapuato,fig.align = "center", out.height='90%', out.width = '90%'}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_irapuato, fig.align = "center", out.height='90%', out.width = '90%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_irapuato, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::
</body>
</html>


### [**San Miguel Allende**]{style="color:#990000"}

<html>
<head>

</head>
<body>
::: wrapper
::: {.box .a}
```{r hotspots_dengue_allende}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "11", 
                                 locality = "San Miguel Allende")

# Step 2. Extract the hotspots ####
#hotspots <- hotspotsdata::den_hotspots[loc, ]

hotspots <- hotspotsdata::hotspots_hex[loc, ]

# Step 3. map
pal <- rcartocolor::carto_pal(n = max(hotspots$intensity_gi), name = "OrYel")
pal_fn <- grDevices::colorRampPalette(pal)
mapgl::maplibre_view(hotspots,
                     column = "intensity_gi",
                     palette = pal_fn) |>
    mapgl::add_continuous_legend(legend_title = "Intensidad",
                          values = seq(from = 1, 
                                       to = max(hotspots$intensity_gi),
                                       by = 1),
                          colors = pal,
                          position = "top-left",
                          width = "250px",
                          draggable = TRUE,
                          style = mapgl::legend_style(element_border_width =0,
                                                      shadow_size = 0,
                                                      shadow = FALSE),
                          interactive = TRUE) |>
    mapgl::add_fullscreen_control(position = "top-left") |> 
    mapgl::add_navigation_control() |>
    mapgl::add_globe_control() |>
    mapgl::add_scale_control()

```
:::

::: {.box .b}
```{r forest_plot_allende, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_allende,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto) +
                  ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                                 axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")))  |> 
  plotly::layout(showlegend = FALSE)






```
:::

::: {.box .d}
```{r gauge_blocks_allende,fig.align = "center", out.height='90%', out.width = '90%'}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_allende, fig.align = "center", out.height='90%', out.width = '90%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_allende, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::
</body>
</html>

### [**Silado**]{style="color:#990000"}

<html>
<head>

</head>
<body>
::: wrapper
::: {.box .a}
```{r hotspots_dengue_silao}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "11", 
                                 locality = "Silao de la Victoria")

# Step 2. Extract the hotspots ####
#hotspots <- hotspotsdata::den_hotspots[loc, ]

hotspots <- hotspotsdata::hotspots_hex[loc, ]

# Step 3. map
pal <- rcartocolor::carto_pal(n = max(hotspots$intensity_gi), name = "OrYel")
pal_fn <- grDevices::colorRampPalette(pal)
mapgl::maplibre_view(hotspots,
                     column = "intensity_gi",
                     palette = pal_fn) |>
    mapgl::add_continuous_legend(legend_title = "Intensidad",
                          values = seq(from = 1, 
                                       to = max(hotspots$intensity_gi),
                                       by = 1),
                          colors = pal,
                          position = "top-left",
                          width = "250px",
                          draggable = TRUE,
                          style = mapgl::legend_style(element_border_width =0,
                                                      shadow_size = 0,
                                                      shadow = FALSE),
                          interactive = TRUE) |>
    mapgl::add_fullscreen_control(position = "top-left") |> 
    mapgl::add_navigation_control() |>
    mapgl::add_globe_control() |>
    mapgl::add_scale_control()

```
:::

::: {.box .b}
```{r forest_plot_silao, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_silao,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto) +
                  ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                                 axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")))  |> 
  plotly::layout(showlegend = FALSE)






```
:::

::: {.box .d}
```{r gauge_blocks_silo,fig.align = "center", out.height='90%', out.width = '90%'}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_silao, fig.align = "center", out.height='90%', out.width = '90%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_silao, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::
</body>
</html>



### [**Salamanca**]{style="color:#990000"}

<html>
<head>

</head>
<body>
::: wrapper
::: {.box .a}
```{r hotspots_dengue_salamanca}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "11", 
                                 locality = "Salamanca")

# Step 2. Extract the hotspots ####
#hotspots <- hotspotsdata::den_hotspots[loc, ]

hotspots <- hotspotsdata::hotspots_hex[loc, ]

# Step 3. map
pal <- rcartocolor::carto_pal(n = max(hotspots$intensity_gi), name = "OrYel")
pal_fn <- grDevices::colorRampPalette(pal)
mapgl::maplibre_view(hotspots,
                     column = "intensity_gi",
                     palette = pal_fn) |>
    mapgl::add_continuous_legend(legend_title = "Intensidad",
                          values = seq(from = 1, 
                                       to = max(hotspots$intensity_gi),
                                       by = 1),
                          colors = pal,
                          position = "top-left",
                          width = "250px",
                          draggable = TRUE,
                          style = mapgl::legend_style(element_border_width =0,
                                                      shadow_size = 0,
                                                      shadow = FALSE),
                          interactive = TRUE) |>
    mapgl::add_fullscreen_control(position = "top-left") |> 
    mapgl::add_navigation_control() |>
    mapgl::add_globe_control() |>
    mapgl::add_scale_control()

```
:::

::: {.box .b}
```{r forest_plot_salamanca, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_salamanca,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto) +
                  ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                                 axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")))  |> 
  plotly::layout(showlegend = FALSE)






```
:::

::: {.box .d}
```{r gauge_blocks_salamanca,fig.align = "center", out.height='90%', out.width = '90%'}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_salamanca, fig.align = "center", out.height='90%', out.width = '90%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_salamanca, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::
</body>
</html>


### [**Acambaro**]{style="color:#990000"}

<html>
<head>

</head>
<body>
::: wrapper
::: {.box .a}
```{r hotspots_dengue_acambaro}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "11", 
                                 locality = "Acambaro")

# Step 2. Extract the hotspots ####
#hotspots <- hotspotsdata::den_hotspots[loc, ]

hotspots <- hotspotsdata::hotspots_hex[loc, ]

# Step 3. map
pal <- rcartocolor::carto_pal(n = max(hotspots$intensity_gi), name = "OrYel")
pal_fn <- grDevices::colorRampPalette(pal)
mapgl::maplibre_view(hotspots,
                     column = "intensity_gi",
                     palette = pal_fn) |>
    mapgl::add_continuous_legend(legend_title = "Intensidad",
                          values = seq(from = 1, 
                                       to = max(hotspots$intensity_gi),
                                       by = 1),
                          colors = pal,
                          position = "top-left",
                          width = "250px",
                          draggable = TRUE,
                          style = mapgl::legend_style(element_border_width =0,
                                                      shadow_size = 0,
                                                      shadow = FALSE),
                          interactive = TRUE) |>
    mapgl::add_fullscreen_control(position = "top-left") |> 
    mapgl::add_navigation_control() |>
    mapgl::add_globe_control() |>
    mapgl::add_scale_control()

```
:::

::: {.box .b}
```{r forest_plot_acambaro, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_acambaro,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto) +
                  ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                                 axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")))  |> 
  plotly::layout(showlegend = FALSE)






```
:::

::: {.box .d}
```{r gauge_blocks_acambaro,fig.align = "center", out.height='90%', out.width = '90%'}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_acambaro, fig.align = "center", out.height='90%', out.width = '90%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_acambaro, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::
</body>
</html>


:::

# [**Hotspots del Vector**]{style="font-size:2pt"}
 
::: panel-tabset

### [**León**]{style="color:#990000"}

```{r vector_hotspots_leon, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_11_guanajuato,
                           locality = "León de Aldama",
                           cve_edo = "11",
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark",
                           static_map = FALSE)
```


### [**Guanajuato**]{style="color:#990000"}

```{r vector_hotspots_guanjuato, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_11_guanajuato,
                           locality = "Guanajuato",
                           cve_edo = "11",
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark",
                           static_map = FALSE)
```


### [**Celaya**]{style="color:#990000"}

```{r vector_hotspots_celaya, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_11_guanajuato,
                           locality = "Celaya",
                           cve_edo = "11",
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark",
                           static_map = FALSE)
```


### [**Irapuato**]{style="color:#990000"}

```{r vector_hotspots_irapuato, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_11_guanajuato,
                           locality = "Irapuato",
                           cve_edo = "11",
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark",
                           static_map = FALSE)
```


### [**San Migueal Allende**]{style="color:#990000"}

```{r vector_hotspots_allende, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_11_guanajuato,
                           locality = "San Miguel Allende",
                           cve_edo = "11",
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark",
                           static_map = FALSE)
```


### [**Silao**]{style="color:#990000"}

```{r vector_hotspots_silao, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_11_guanajuato,
                           locality = "Silao de la Victoria",
                           cve_edo = "11",
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark",
                           static_map = FALSE)
```


### [**Salamanca**]{style="color:#990000"}

```{r vector_hotspots_salamanca, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_11_guanajuato,
                           locality = "Salamanca",
                           cve_edo = "11",
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark",
                           static_map = FALSE)
```


### [**Acambaro**]{style="color:#990000"}

```{r vector_hotspots_acambaro, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_11_guanajuato,
                           locality = "Acambaro",
                           cve_edo = "11",
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark",
                           static_map = FALSE)
```



:::


# [**Escenarios Epidemiológicos**]{style="font-size:2pt"}
 
::: panel-tabset

### [**Leon**]{style="color:#990000"}

```{r risk_ageb_leon}
# Step 1. calculate the risk ####
denhotspots::risk_ageb(betas = denmex::eggs_betas_11_guanajuato,
                       hotspots = denmex::den_hotspots,
                       intensity_perc = 25,
                       locality = "Leon de Aldama",
                       cve_edo = "11") |>
  denhotspots::map_risk(staticmap = FALSE)

```


### [**Guanajuato**]{style="color:#990000"}

### [**Irapuato**]{style="color:#990000"}

### [**San Miguel Allende**]{style="color:#990000"}

### [**Silao**]{style="color:#990000"}


### [**Salamanca**]{style="color:#990000"}


### [**Acambaro**]{style="color:#990000"}


### [**Celaya**]{style="color:#990000"}


:::

